Topology and geometry of the dark matter web: a multistream view

### 2018-03-07 Draft/Chapter4/Source_v2/draft_paper1.tex
initial commit

\begin{comment}
\documentclass[fleqn,usenatbib,useAMS]{mnras}
%\documentclass[referee,fleqn,usenatbib,useAMS]{mnras}

@@

% We also introduce, for the first time, a framework to detect dark matter haloes in multistream fields. Closed compact regions hosting local maxima of the multistream field are detected using local geometrical conditions and properties of the Lagrangian sub-manifold. All the halo particles are guaranteed to be completely outside void regions of the Universe. Majority of the halo candidates are embedded in the largest structure that percolates the entire volume.

@@

\begingroup
\let\clearpage\relax
%\tableofcontents
\endgroup
\newpage
\end{comment}
%-------------------------------------------------------------------------------------------------------------------------------
\chapter{Topology and geometry of the dark matter web: a multistream view}

@@

\begin{table}
  \caption{Parameters for the simulation boxes: Side length $L$, number of particles $N_p$, mass of each particle $m_p$, and the gravitational softening length $\epsilon$ for the GADGET simulations are shown.}
\begin{tabular}{l|c|c|c| }
\hline
$L$ & $N_p$ & $m_p$ & $\epsilon$  \\  \hline
$100 h^{-1} Mpc$ & $128^3$      & $ 3.65 \times 10^{10} h^{-1} M_{\odot}$  & $20 h^{-1} kpc$\\ \hline
$100 h^{-1} Mpc$ & $256^3$      & $ 4.57 \times 10^{9} h^{-1} M_{\odot}$  & $10 h^{-1} kpc$ \\ \hline

@@

\begin{figure} 
\centering\includegraphics[width=8cm]{Chapter4/Source_v2/fig1.pdf} 
\caption{3D rendering of the multistream field: the cosmic web structure of a $ 50 h^{-1} \text{Mpc} \times 50 h^{-1} \text{Mpc} \times 50 h^{-1} \text{Mpc}$ slice in a simulation box of side length $100 h^{-1}$ Mpc and $128^3$ particles. The multistream field is calculated at 8 times the native resolution. void(black) is a percolating structure with $n_{str} = 1$. Regions $n_{str} \geq 17$ show a filamentary structure (gray) and the bright spots at the intersections of the filaments are regions with $n_{str} \geq 100$. }
\label{fig:full}
\end{figure}

% \bmath changed to \bf --- check if it's alright. 
The multistream field objectively characterizes the level of non-linearity in the cosmic web. The `number-of-streams' field or $n_{str}({\bmath x})$ is computed from the Lagrangian sub-manifold ${\bmath x}({\bmath q})$, which is a continuous three-dimensional sheet in a six-dimensional  $({\bmath q}, {\bmath x})$ space. In this paper, we utilize the tessellation implementation by \cite{Shandarin2012} to calculate the multistream flow field on the GADGET-2 snapshot at $z=0$. This implementation only requires initial and final coordinates of the dark matter particles.

@@

\begin{figure*}
\begin{minipage}[t]{0.99\linewidth}
 \centering\includegraphics[height=11cm]{Chapter4/Source_v2/fig2.pdf} 
\end{minipage}\hfill
\caption{Opposite faces of the multistream field for the simulation box with $N_p = 128^3$. Non-void regions (gray) have $n_{str} > 1$. The largest void (white) in the entire field spans over the entire box. Rest of the smaller isolated voids (red) occupy very small volume fraction. }
\label{fig:voidFace}
\end{figure*}

@@

\begin{figure}
\begin{minipage}[t]{.99\linewidth}
\centering\includegraphics[width=8.cm]{Chapter4/Source_v2/fig3.pdf} 
\end{minipage}\hfill
\caption{Single-streaming void distribution on diagnostic spheres around FOF-haloes are considered. At radius $r_{10}$, each diagnostic sphere has $n_{str} = 1$ on 10 per cent of its spherical surface. Distribution function of $r_{10}$ (blue) and FOF-radii $r_{vir}$ (red) are shown. Inner plot shows the distribution function of $r_{10}/r_{vir}$. The haloes within the dashed line have at least 10 per cent of their virial-surfaces in contact with $n_{str} = 1$ regions.}
\label{fig:Vfr_all}
\end{figure}

@@

\begin{figure}
\begin{minipage}[t]{.99\linewidth}
  \centering\includegraphics[width=8.cm]{Chapter4/Source_v2/fig4.pdf} 
\end{minipage}\hfill
\caption{Percolation plot in the multistream field and mass density. Two density estimators - CIC and DTFE are shown. Percolation transition (at $f_1/f_{ES} = 0.5$ shown by the horizontal red line) occurs at smaller excursion set volumes for the multistream field, as seen by the dashed lines for both the curves. It is worth stressing that the percolation curves for $n_{str}$ field are bounded by conditions $f_{ES} < 0.1$. }
\label{fig:PercTh}
\end{figure}

\begin{figure*}
\begin{minipage}[t]{.99\linewidth}
  \centering\includegraphics[height=5.7cm]{Chapter4/Source_v2/fig5.pdf} 
\end{minipage}\hfill
\caption{Percolation threshold in the multistream (left panel) and matter density fields. Matter density is calculated using DTFE (middle panel) and CIC (right panel) with a refinement factor of 2 in the simulation with $256^3$ particles. The volume fraction of excursion set and the filling fraction of the largest structure is shown. Percolation transition in multistream field at $n_{str} = 17$ is shows by the dashed vertical line. Percolation at $\rho_{DTFE}/ \rho_b = 5.16 $ and $\rho_{CIC}/ \rho_b = 5.49 $ are shown by the dashed vertical line. }
\label{fig:HaloFil}
\end{figure*}

@@

\section{Local geometry of the multistream field}
\label{sec:hessian}
\begin{figure}
\begin{minipage}[t]{.99\linewidth}
  \centering\includegraphics[width=8.cm]{Chapter4/Source_v2/fig6.pdf}

@@

\begin{figure}
\begin{minipage}[t]{.99\linewidth}
  \centering\includegraphics[width=8.cm]{Chapter4/Source_v2/fig7.pdf}

@@

\begin{figure*}
\begin{minipage}[t]{0.99\linewidth}
 \centering\includegraphics[height=14.cm]{Chapter4/Source_v2/fig8.pdf} 
\end{minipage}\hfill
\caption{Eigenvalues of the Hessian matrix $\mathbfss{H}(-n_{str})$ in a slice of 50 $h^{-1}$ Mpc $\times$ 50 $h^{-1}$ Mpc slice of the simulation box of $128^3$ particles. Variation in the eigenvalues in the multistreaming web structure is shown. The largest eigenvalue $\lambda_1$ (top left panel) has positive values throughout the structure. The smallest eigenvalue $\lambda_3$ (bottom left) has negative values surrounding positive definite regions of the $n_{str}$ field. Corresponding multistream field is shown in the bottom right panel for single, three and more than five streams.}
\label{fig:evals123}
\end{figure*}

\begin{figure}
\begin{minipage}[t]{.99\linewidth}
  \centering\includegraphics[width=6.cm]{Chapter4/Source_v2/fig9.pdf} 
\end{minipage}\hfill
\caption{Surfaces identified in the multistream field. Blue regions are closed regions with $\lambda_3 > 0 $, which we identify as two haloes. Other surface has an open curvature along one direction, with $\lambda_1 > \lambda_2 > 0$ and $ \lambda_3 < 0$. }
\label{fig:SmallBox}
\end{figure}

@@

\begin{figure}
\begin{minipage}[t]{.99\linewidth}
  \centering\includegraphics[width=8.cm]{Chapter4/Source_v2/fig10.pdf}

@@

\begin{figure}
\begin{minipage}[t]{.99\linewidth}
  \centering\includegraphics[width=8.cm]{Chapter4/Source_v2/fig11.pdf}

@@

\begin{figure}
\begin{minipage}[t]{.99\linewidth}
  \centering\includegraphics[width=10.cm]{Chapter4/Source_v2/fig12a.pdf}
  \centering\includegraphics[width=10.cm]{Chapter4/Source_v2/fig12b.pdf}  
\end{minipage}\hfill
\caption{Two top panels show  three contours ($n_{str}=3, 11, 17$) in a slice $100h^{-1} \text{Mpc} \times 100h^{-1} \text{Mpc} \times 10h^{-1} \text{Mpc}$ in the simulation with $128^3$ particles, computed at two refinement factors: 2(upper) and 8 (lower). The bottom panel shows the caustic surfaces in the same slice. }
\label{fig:NstrCaust}
\end{figure}

@@

% Everything commented below - including Appendix that could be added later.

@@

\begin{figure}
\begin{minipage}[t]{.99\linewidth}
  \centering\includegraphics[width=8.cm]{Chapter4/Source_v2/fig13.pdf} 
\end{minipage}\hfill
\caption{ Multi-streaming in one-dimension gravitational collapse. Top panel: $(\bmath{p}, \bmath{x})$ phase-space representation redshift $z_{ini}$ and $z = 0$. Dots represent the dark matter particles. Initially the mass particles are in the linear stage of evolution. At $z = 0$, multiple values of $\bmath{p}(\bmath{x})$ is seen in the collapsed regions. Middle panel: Equivalent Lagrangian sub-manifold $\bmath{q}(\bmath{x})$. At $z_{ini}$, the dashed line represents $\bmath{q} = \bmath{x}$. Number of streams are parametrized from this sub-manifold. Bottom panel: The multistream field $n_{str}$ and the number-density using CIC algorithm, $n_{CIC}$ at $z = 0$. }
\label{fig:phase1d}
\end{figure}

@@

% An illustration of the projection is shown in \autoref{fig:check1d} for a simpler function $f(x)$ in one-dimensional domain. The eigenvalue criteria for regions are simplified: for instance, $\frac{\partial^2 f}{\partial x^2} < 0$ for convex region. Projection of these regions onto coordinate space is shown in the shaded regions. This is different from regions within a contour, which is the projection of the curve along which the function has a constant value. Boundaries of these two regions may, but not necessarily, intersect.    
% 
% \begin{figure}
% \begin{minipage}[t]{.99\linewidth}
%   \centering\includegraphics[width=8.cm]{fig25.pdf} 
% \end{minipage}\hfill
% \caption{Projections of regions of $f(x)$ from (1+1)-dimensional function space onto one-dimensional coordinate space. Convex regions and regions above a threshold of an arbitrary function $f(x)$ are shown. Both the regions intersect around a few maxima, but not universally.}
% \label{fig:check1d}
% \end{figure}

@@

\end{document}
\end{comment}

>>>>

\documentclass[fleqn,usenatbib,useAMS]{mnras}
%\documentclass[referee,fleqn,usenatbib,useAMS]{mnras}

@@

\begin{comment}
We also introduce, for the first time, a framework to detect dark matter haloes in multistream fields. Closed compact regions hosting local maxima of the multistream field are detected using local geometrical conditions and properties of the Lagrangian sub-manifold. All the halo particles are guaranteed to be completely outside void regions of the Universe. Majority of the halo candidates are embedded in the largest structure that percolates the entire volume.
\end{comment}

@@

\begingroup
\let\clearpage\relax
%\tableofcontents
\endgroup
\newpage
%-------------------------------------------------------------------------------------------------------------------------------

@@

\begin{table}
  \caption{Parameters for the simulation boxes: Side length $L$, number of particles $N_p$, mass of each particle $m_p$, and the gravitational softening length $\epsilon$ for the GADGET simulations are shown.}
\begin{tabular}{l|c|c|c| }
\hline
$L$ & $N_p$ & $m_p$ & $\epsilon$  \\  \hline
$100 h^{-1} Mpc$ & $128^3$      & $ 3.65 \times 10^{10} h^{-1} M_{\sun}$  & $20 h^{-1} kpc$\\ \hline
$100 h^{-1} Mpc$ & $256^3$      & $ 4.57 \times 10^{9} h^{-1} M_{\sun}$  & $10 h^{-1} kpc$ \\ \hline

@@

\begin{figure} 
\centering\includegraphics[width=8cm]{fig1.pdf} 
\caption{3D rendering of the multistream field: the cosmic web structure of a $ 50 h^{-1} \text{Mpc} \times 50 h^{-1} \text{Mpc} \times 50 h^{-1} \text{Mpc}$ slice in a simulation box of side length $100 h^{-1}$ Mpc and $128^3$ particles. The multistream field is calculated at 8 times the native resolution. void(black) is a percolating structure with $n_{str} = 1$. Regions $n_{str} \geq 17$ show a filamentary structure (gray) and the bright spots at the intersections of the filaments are regions with $n_{str} \geq 100$. }
\label{fig:full}
\end{figure}

The multistream field objectively characterizes the level of non-linearity in the cosmic web. The `number-of-streams' field or $n_{str}(\bmath{x})$ is computed from the Lagrangian sub-manifold $\bmath{x}(\bmath{q})$, which is a continuous three-dimensional sheet in a six-dimensional  $(\bmath{q}, \bmath{x})$ space. In this paper, we utilize the tessellation implementation by \cite{Shandarin2012} to calculate the multistream flow field on the GADGET-2 snapshot at $z=0$. This implementation only requires initial and final coordinates of the dark matter particles.

@@

\begin{figure*}
\begin{minipage}[t]{0.99\linewidth}
 \centering\includegraphics[height=11cm]{fig2.pdf} 
\end{minipage}\hfill
\caption{Opposite faces of the multistream field for the simulation box with $N_p = 128^3$. Non-void regions (gray) have $n_{str} > 1$. The largest void (white) in the entire field spans over the entire box. Rest of the smaller isolated voids (red) occupy very small volume fraction. }
\label{fig:voidFace}
\end{figure*}

@@

\begin{figure}
\begin{minipage}[t]{.99\linewidth}
\centering\includegraphics[width=8.cm]{fig3.pdf} 
\end{minipage}\hfill
\caption{Single-streaming void distribution on diagnostic spheres around FOF-haloes are considered. At radius $r_{10}$, each diagnostic sphere has $n_{str} = 1$ on 10 per cent of its spherical surface. Distribution function of $r_{10}$ (blue) and FOF-radii $r_{vir}$ (red) are shown. Inner plot shows the distribution function of $r_{10}/r_{vir}$. The haloes within the dashed line have at least 10 per cent of their virial-surfaces in contact with $n_{str} = 1$ regions.}
\label{fig:Vfr_all}
\end{figure}

@@

\begin{figure}
\begin{minipage}[t]{.99\linewidth}
  \centering\includegraphics[width=8.cm]{fig4.pdf} 
\end{minipage}\hfill
\caption{Percolation plot in the multistream field and mass density. Two density estimators - CIC and DTFE are shown. Percolation transition (at $f_1/f_{ES} = 0.5$ shown by the horizontal red line) occurs at smaller excursion set volumes for the multistream field, as seen by the dashed lines for both the curves. It is worth stressing that the percolation curves for $n_{str}$ field are bounded by conditions $f_{ES} < 0.1$. }
\label{fig:PercTh}
\end{figure}

\begin{figure*}
\begin{minipage}[t]{.99\linewidth}
  \centering\includegraphics[height=5.7cm]{fig5.pdf} 
\end{minipage}\hfill
\caption{Percolation threshold in the multistream (left panel) and matter density fields. Matter density is calculated using DTFE (middle panel) and CIC (right panel) with a refinement factor of 2 in the simulation with $256^3$ particles. The volume fraction of excursion set and the filling fraction of the largest structure is shown. Percolation transition in multistream field at $n_{str} = 17$ is shows by the dashed vertical line. Percolation at $\rho_{DTFE}/ \rho_b = 5.16 $ and $\rho_{CIC}/ \rho_b = 5.49 $ are shown by the dashed vertical line. }
\label{fig:HaloFil}
\end{figure*}

@@

\section{Local geometry of the multistream field}
\label{sec:hessian}
\begin{figure}
\begin{minipage}[t]{.99\linewidth}
  \centering\includegraphics[width=8.cm]{fig6.pdf}

@@

\begin{figure}
\begin{minipage}[t]{.99\linewidth}
  \centering\includegraphics[width=8.cm]{fig7.pdf}

@@

\begin{figure*}
\begin{minipage}[t]{0.99\linewidth}
 \centering\includegraphics[height=14.cm]{fig8.pdf} 
\end{minipage}\hfill
\caption{Eigenvalues of the Hessian matrix $\mathbfss{H}(-n_{str})$ in a slice of 50 $h^{-1}$ Mpc $\times$ 50 $h^{-1}$ Mpc slice of the simulation box of $128^3$ particles. Variation in the eigenvalues in the multistreaming web structure is shown. The largest eigenvalue $\lambda_1$ (top left panel) has positive values throughout the structure. The smallest eigenvalue $\lambda_3$ (bottom left) has negative values surrounding positive definite regions of the $n_{str}$ field. Corresponding multistream field is shown in the bottom right panel for single, three and more than five streams.}
\label{fig:evals123}
\end{figure*}

\begin{figure}
\begin{minipage}[t]{.99\linewidth}
  \centering\includegraphics[width=6.cm]{fig9.pdf} 
\end{minipage}\hfill
\caption{Surfaces identified in the multistream field. Blue regions are closed regions with $\lambda_3 > 0 $, which we identify as two haloes. Other surface has an open curvature along one direction, with $\lambda_1 > \lambda_2 > 0$ and $ \lambda_3 < 0$. }
\label{fig:SmallBox}
\end{figure}

@@

\begin{figure}
\begin{minipage}[t]{.99\linewidth}
  \centering\includegraphics[width=8.cm]{fig10.pdf}

@@

\begin{figure}
\begin{minipage}[t]{.99\linewidth}
  \centering\includegraphics[width=8.cm]{fig11.pdf}

@@

\begin{figure}
\begin{minipage}[t]{.99\linewidth}
  \centering\includegraphics[width=10.cm]{fig12a.pdf}
  \centering\includegraphics[width=10.cm]{fig12b.pdf}  
\end{minipage}\hfill
\caption{Two top panels show  three contours ($n_{str}=3, 11, 17$) in a slice $100h^{-1} \text{Mpc} \times 100h^{-1} \text{Mpc} \times 10h^{-1} \text{Mpc}$ in the simulation with $128^3$ particles, computed at two refinement factors: 2(upper) and 8 (lower). The bottom panel shows the caustic surfaces in the same slice. }
\label{fig:NstrCaust}
\end{figure}

@@

@@

\begin{figure}
\begin{minipage}[t]{.99\linewidth}
  \centering\includegraphics[width=8.cm]{fig13.pdf} 
\end{minipage}\hfill
\caption{ Multi-streaming in one-dimension gravitational collapse. Top panel: $(\bmath{p}, \bmath{x})$ phase-space representation redshift $z_{ini}$ and $z = 0$. Dots represent the dark matter particles. Initially the mass particles are in the linear stage of evolution. At $z = 0$, multiple values of $\bmath{p}(\bmath{x})$ is seen in the collapsed regions. Middle panel: Equivalent Lagrangian sub-manifold $\bmath{q}(\bmath{x})$. At $z_{ini}$, the dashed line represents $\bmath{q} = \bmath{x}$. Number of streams are parametrized from this sub-manifold. Bottom panel: The multistream field $n_{str}$ and the number-density using CIC algorithm, $n_{CIC}$ at $z = 0$. }
\label{fig:phase1d}
\end{figure}

@@

\begin{comment}
An illustration of the projection is shown in \autoref{fig:check1d} for a simpler function $f(x)$ in one-dimensional domain. The eigenvalue criteria for regions are simplified: for instance, $\frac{\partial^2 f}{\partial x^2} < 0$ for convex region. Projection of these regions onto coordinate space is shown in the shaded regions. This is different from regions within a contour, which is the projection of the curve along which the function has a constant value. Boundaries of these two regions may, but not necessarily, intersect.

\begin{figure}
\begin{minipage}[t]{.99\linewidth}
  \centering\includegraphics[width=8.cm]{fig25.pdf} 
\end{minipage}\hfill
\caption{Projections of regions of $f(x)$ from (1+1)-dimensional function space onto one-dimensional coordinate space. Convex regions and regions above a threshold of an arbitrary function $f(x)$ are shown. Both the regions intersect around a few maxima, but not universally.}
\label{fig:check1d}
\end{figure}

@@

\end{document}
